---
title: "Lab 01 - Hello R"
author: "YG"
date: "1/23/2023"
output: 
  html_document: 
    fig_width: 6
    fig_height: 3
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

```{r}
dim(datasaurus_dozen)
datasaurus_dozen$dataset<-as.factor(datasaurus_dozen$dataset)
summary(datasaurus_dozen)
table(datasaurus_dozen$dataset)
```

### Exercise 2

First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x,y))
```

### Exercise 3

Plot the star dataset.

```{r plot-star}
star_data<-datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x=x, y=y)) +
  geom_point()
```

Calculate the correlation between x and y for the star dataset.

```{r cor-star}
star_data %>%
  summarize(r = cor(x,y))
```

The dino data set has higher correlation than the star data set.

### Exercise 4

Plot the circle dataset.

```{r plot-circle}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x=x, y=y)) +
  geom_point()
```

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x,y))
```

The circle data set has a higher correlation than the dino data set.

### Exercise 5

```{r plot the entire data set, fig.height=8, fig.width=6}
ggplot(datasaurus_dozen, aes(x=x, y=y, color = dataset)) + 
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```

```{r correlation of all the dataset}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x,y)) %>%
  print(13)
```

Here are some helpful tips :)

    filter() is for extracting rows

    group_by() is for grouping datasets by assigned column

    ungroup() cancels the grouping

    summarize() is often used with group_by(). This function can print the output according to the group_by().

    facet_grid(y~x,…) creates a grid with variable y as a row, variable x as a column

    facet_wrap(x,… ) is useful when there is only one variable

