---
title: "Data and Visualization"
author: "YG"
date: "2023-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
options(tibble.print_min = 10)
heights <- read.csv("data/raw-data/highest-points-by-state.csv")
heights$elevation <- heights$elevation *.3048
```
Order the data by elevation
```{r}
arrange(heights, elevation)
arrange(heights, desc(elevation))
```

Stem plot
```{r}
stem(heights$elevation)
```
density estimate
```{r}
ggplot(heights, aes(x=elevation)) + geom_density() + geom_rug()
```
Explore Dataset

```{r}
#glimpse(starwars)
#nrow(starwars)
#ncol(starwars)
#dim(starwars)
#summary(starwars)
#head(starwars)
#tail(starwars)
#names(starwars)
#length(starwars)
 
plot(mass~height, starwars)
```

Plot Mass vs.height
```{r}
ggplot(data = starwars, mapping = aes(x=height, y=mass)) + geom_point() +
  labs(title = "Mass vs. height of starwar characters",
       x = "Height (cm)", y="Weight (kg)")
```
ggplot
```{r}
library(palmerpenguins)
glimpse(penguins)
```

```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species,
                                      shape=species,
                                      size=body_mass_g,
                                      alpha=flipper_length_mm)) +
  geom_point() +
  labs(title = "Bill Depth and Length",
       subtitle = "Dimentions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", y="Bill length (mm)",
       caption = "Source: Palmer Station LTER/palmerpenguins package") +
  scale_color_viridis_d()
```

```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species,
                                      shape=species)) +
  geom_point(size=2, alpha=0.5) +
  labs(title = "Bill Depth and Length",
       subtitle = "Dimentions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", y="Bill length (mm)",
       caption = "Source: Palmer Station LTER/palmerpenguins package") +
  scale_color_viridis_d()
```
Mapping: determine the size, alpha, etc. of points based on the values of a variable in the data. It goes into aes()
Setting: determine the size, alpha, etc. of points NOT based on the values of a variable in the data. It goes into geom_*()

Facet
```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species)) +
  geom_point() +
  facet_grid(species~island)
```
```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species)) +
  geom_point() +
  facet_grid(species~sex)
```


```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species)) +
  geom_point() +
  facet_wrap(species~island)
```


```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species)) +
  geom_point() +
  facet_wrap(~species)
```

```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species)) +
  geom_point() +
  facet_grid(.~species)
```

```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=bill_length_mm, 
                                      color=species)) +
  geom_point() +
  facet_wrap(~species, ncol=2)
```
Faceting Summary

facet_grid: 2d grid, rows~columns, use . for no split
facet_wrap: 1d ribbon wrapped according to number of rows and columns specified or available in the plotting area.


Visualizing numerical data
univariate data: distribution of a single variable
bivariate data: the relationship between two variables
multivariate data: relationship between many variables at once. Usually focuing on the relationship between two while conditioning for others.

```{r}
library(openintro)
glimpse(loans_full_schema)
```
```{r}
loans<-loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```
Describing the shapes of numerical data
1. shape: skewness, symmetric; modality (unimodal, bimodal, multimodal, uniform)
2. center: mean, median, mode
3. spread: range, standard deviation, inter-quartile range(IQR)
4. unusual observations

```{r}
ggplot(loans, aes(x=loan_amount)) +
  geom_histogram()
```


```{r}
ggplot(loans, aes(x=loan_amount)) +
  geom_histogram(binwidth = 1000)
```


```{r}
ggplot(loans, aes(x=loan_amount)) +
  geom_histogram(binwidth = 20000)
```

```{r}
ggplot(loans, aes(x=loan_amount, fill=homeownership)) +
  geom_histogram(binwidth = 5000,
                 alpha=0.5) +
  labs(x="loan amount", y="frequency", title = "Amounts of lending club loans")
```

```{r}
ggplot(loans, aes(x=loan_amount)) +
geom_density()
```

```{r}
ggplot(loans, aes(x=loan_amount)) +
geom_density(adjust = 0.5) #default adjust = 1
```
```{r}
ggplot(loans, aes(x=loan_amount)) +
geom_density(adjust = 2)
```
```{r}
ggplot(loans, aes(x=loan_amount, fill = homeownership)) +
geom_density(adjust=2, alpha = 0.5) +
  labs(x="loan amount ($)", y="density", title="Amounts of lending club loans",
       fill="home ownership")
```
```{r}
ggplot(loans, aes(x=interest_rate)) + 
  geom_boxplot()
```
```{r}
ggplot(loans, aes(x=interest_rate)) + 
  geom_boxplot() +
  labs(x="interest rate (%)", y=NULL, title="Interest rate of Lending club loans") +
  theme(axis.ticks.y=element_blank(),
        axis.text.y=element_blank())
```
```{r}
ggplot(loans, aes(x=interest_rate,
                  y=grade, color=grade)) + 
  geom_boxplot() +
  labs(x="interest rate (%)", y=NULL, title="Interest rate of Lending club loans",
       subtitle = "by grade of loan") +
  theme(axis.ticks.y=element_blank(),
        axis.text.y=element_blank())
```
Relationships numeric variables

```{r}
ggplot(loans, aes(x=debt_to_income, y=interest_rate)) +
  geom_point()
```
```{r}
ggplot(loans, aes(x=debt_to_income, y=interest_rate)) +
  geom_hex()
```
```{r}
ggplot(loans %>% filter(debt_to_income<100), aes(x=debt_to_income, y=interest_rate)) +
  geom_hex()
```
Visualizing categorical data

```{r}
starwars <- starwars %>%
  mutate(hair_color2 = 
           fct_other(hair_color,
                     keep=c("black",'brown','blond')))
```

```{r}
ggplot(starwars, aes(x=gender)) +
  geom_bar()
```
```{r}
ggplot(starwars, aes(x=gender, fill=hair_color)) +
  geom_bar()
```

```{r}
ggplot(starwars, aes(x=gender, fill=hair_color2)) +
  geom_bar() +
  coord_flip()
```

```{r}
ggplot(starwars, aes(x=gender, fill=hair_color2)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(x="proportion", y="gender", fill="hair color",
       title = "hair color of starwars", subtitle = "by gender")
```
```{r}
ggplot(loans, aes(x=homeownership, y=loan_amount)) +
  geom_violin()
```
```{r}
library(ggridges)
ggplot(loans, aes(x=loan_amount, y=grade, fill=grade)) +
  geom_density_ridges(alpha=0.5)
```
Basic care and feeding data to R

rm(list=ls()) #console command to clear the environment
