---
title: "Tidyverse"
author: "YG"
date: "2023-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Tidy data
1. each variable forms a colum
2. each observation forms a row
3. each type of observational unit forms a table

```{r}
starwars %>%
  select(name, height, mass)
```

```{r}
starwars %>%
  group_by(gender) %>%
  summarize(avg_ht=mean(height, na.rm = T))
```

#Data Structures in R

Matrix: requires all the elements to be of the same type
Data frame: allows for mixed types of variables
Tibble: like a data frame, but works more consistently

A grammar of data wrangling

select: pick columns by name
arrange: reorder rows
slice: pick rows using indexes
filter: pick rows matching criteria
distinct: filter for unique rows
mutate: add new variables
summarize: reduce variables to values
group_by: for grouped operations

#Rules of dplyr
first argument is always a data frame
subsequent argument say what to do with that data frame
always return a data frame
don't modify in place

##read in data
```{r}
h1<-read_csv("H1.csv", show_col_types = FALSE)
h1<- h1 %>%
  mutate(hotel="h1")
h2<-read_csv("H2.csv", show_col_types = FALSE)
h2<-h2 %>%
  mutate(hotel="h2")
hotels<-rbind(h1,h2)
#names(h1)
#names(h2)
```

```{r}
names(hotels)
#glimpse(hotels)
```
```{r}
select(hotels, hotel, LeadTime)
```
```{r}
hotels %>%
  select(hotel, LeadTime) %>%
  arrange(desc(LeadTime))
```
##Pipes
It is a technique for passing information from one process to another.
%>% used mainly in dplyr (Cmd+Shift+M)
+ used in ggplots
code styling (always a space before the pipe sign, and a line break after the pipe sign)

Use the dot to
1. send results to a function argument other than first one or
2. use the previous result for multiple arguments

```{r}
hotels %>%
  filter(hotel == 'h1') %>%
  lm(ADR~LeadTime, data=.)
```

#Introduction to dplyr

##load library
```{r}
library(tidyverse)
library(gapminder)
```

```{r}
gapminder
```

```{r}
as_tibble(iris)
```
###Think before you create exerpts of your data

###use filter() to subset data row-wise
```{r}
filter(gapminder, lifeExp<29)
filter(gapminder, country == "Rwanda", year>1979)
filter(gapminder, country %in% c("Rwanda","Afghanistan"))

#base R
gapminder[gapminder$lifeExp<29,]
subset(gapminder, country =="canada")
```
###use select to subset the data on variables or columns
```{r}
select(gapminder, year, lifeExp)

gapminder %>% 
  select(year, lifeExp) %>% 
  head(4)

#base R
gapminder[gapminder$country == "Cambodia", c("year", "lifeExp")]
```
#hands on data wrangling

###select() examples
starts_with(): starts with a prefix
ends_with(): ends with a prefix
contains(): contains a literal string
num_range(): matches a numerical range
one_of(): matches variable names in a character vector
everything(): matches all variables
last_col(): select last variable, possibly with an offset
matches(): matches a regular expression

```{r}
hotels %>% 
  select(-hotel) #exclude certain columns

hotels %>% 
  select(hotel:ArrivalDateMonth) #select a consective columns

hotels %>% 
  select(starts_with('Arrival')) #select variables with certain characteristics
  
hotels %>% 
  select(ends_with('Arrival')) 

hotels %>% 
  select(starts_with('Month'))

```
###arrange() examples
```{r}
hotels %>% 
  select(Adults, Children, Babies) %>% 
  arrange(Babies)

hotels %>% 
  select(Adults, Children, Babies) %>% 
  arrange(desc(Babies))
```

###slice() examples
```{r}
hotels %>% 
  slice(1:5) #slice for certain row numbers

last_row<-nrow(hotels)
hotels %>% 
  slice((last_row-4):last_row) #get the last 5 rows
```

###filter() examples

```{r}
hotels %>% 
  filter(
    Adults == 0,
    Children >= 1
  ) %>% 
  select(Adults, Children, Babies)

hotels %>% 
  filter(
    Adults == 0,
    Children >= 1 | Babies >1
  ) %>% 
  select(Adults, Children, Babies)
```

###distinct() examples
```{r}
hotels %>% 
  distinct(MarketSegment) %>% 
  arrange(MarketSegment)

hotels %>% 
  distinct(hotel, MarketSegment) %>% 
  arrange(hotel, MarketSegment)
```

###count() examples
```{r}
hotels %>% 
  count(MarketSegment)

hotels %>% 
  count(MarketSegment, sort = TRUE) #sort based on the frequency count

hotels %>% 
  count(MarketSegment) %>% 
  arrange(n)

hotels %>% 
  count(MarketSegment) %>% 
  arrange(desc(n))

hotels %>% 
  count(hotel, MarketSegment)

hotels %>% 
  count(MarketSegment, hotel)
```
###mutate examples

```{r}
hotels %>% 
  mutate(little_one = Children + Babies) %>% 
  select(Children, Babies, little_one) %>% 
  arrange(desc(little_one))

hotels %>% 
  mutate(little_one = Children + Babies) %>% 
  filter(
    little_one >= 1,
    hotel == 'h1'
  ) %>% 
  select(hotel, little_one)

hotels %>% 
  mutate(little_one = Children + Babies) %>% 
  count(hotel, little_one) %>% 
  mutate(prop = n/sum(n))
```
###summarize examples

```{r}
hotels %>% 
  group_by(hotel) %>% 
  summarize(mean_adr = mean(ADR), 
            min_adr = min(ADR),
            n=n())

hotels %>% 
  count(hotel)
```
###rename examples newname = currentname

```{r}
hotels %>% 
  rename(HotelType = hotel)
```
####select() can rename and reposition variables
```{r}
hotels %>% 
  filter(hotel == "h1", IsCanceled == "0") %>% 
  select(lt=LeadTime, ArrivalDateYear, ArrivalDateMonth) %>% 
  select(ArrivalDateYear, everything())
```
```{r}
library(gapminder)
my_gap<-gapminder
```

#####Count
```{r}
my_gap %>% 
  group_by(continent) %>% 
  tally()
```

```{r}
my_gap %>% 
  group_by(continent) %>% 
  summarize(n=n(),
            n_countries = n_distinct(country))
```
#####summarize_at()
```{r}
my_gap %>% 
  filter(year %in% c(1952, 2007)) %>% 
  group_by(continent) %>% 
  summarize_at(vars(lifeExp, gdpPercap), list(~mean(.), ~median(.)))
```
```{r}
#first() function extracts the first value from a vector. Notice that first() is operating on the vector of life expectancies within each country group
my_gap %>% 
  group_by(country) %>% 
  select(country, year, lifeExp) %>% 
  mutate(lifeExp_gain = lifeExp - first(lifeExp)) %>% 
  filter(year<1963)
```

#####window functions
```{r}
my_gap %>% 
  filter(continent == "Asia") %>% 
  select(year, country, lifeExp) %>% 
  group_by(year) %>% 
  filter(min_rank(desc(lifeExp))<2 | min_rank(lifeExp)<2) %>% 
  arrange(year) %>% 
  print(n=Inf)
```
```{r}
my_gap %>% 
  filter(continent == "Asia") %>% 
  select(year, country, lifeExp) %>% 
  group_by(year) %>% 
  mutate(le_rank = min_rank(lifeExp),
         le_desc_rank = min_rank(desc(lifeExp))) %>% 
  filter(country %in% c("Afghanistan", "Japan", "Thailand"))
```
#####top_n()

```{r}
my_gap %>% 
  filter(continent == "Asia") %>% 
  select(year, country, lifeExp) %>% 
  arrange(year) %>% 
  group_by(year) %>% 
  top_n(1, wt=lifeExp) #gets the min
#top_n(1, wt=desc(lifeExp)) #gets the max
```

```{r}
my_gap %>% 
  select(country, year, continent, lifeExp) %>% 
  group_by(continent, country) %>% 
  mutate(le_delta=lifeExp - lag(lifeExp)) %>% 
  summarize(worst_le_delta = min(le_delta, na.rm=TRUE)) %>% 
  top_n(-1, wt = worst_le_delta) %>% #within continent, retain the row with the lowest worst_le_delta
  arrange(worst_le_delta)
```

#Working with multiple data frames

##join data frames
something_join(x,y)
left_join():all the rows from x
right_join():all the rows from y
full_join():all the rows from x and y
semi_join():all the rows from x where there are matching values in y, keeping just columns from x
inner_join():all the rows from x where there are matching values in ye, return all combination of multiple matches in the case of multiple matches
anti_join():all the rows from x where there are no matching values in y, never deplicate rows of x

```{r}
profession %>% 
  left_join(dates, by="id")
```



