---
title: "DataTypes"
author: "YG"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# cat_lovers %>% 
#   mutate(number_of_cats = case_when(
#     name == "Ginger Clark" ~"2",
#     name == "Doug Bass"    ~"3",
#     TRUE                   ~number_of_cats
#   ),
#   number_of_cats = as.numeric(number_of_cats)) %>% 
#   summarize(mean_cats = mean(number_of_cats))
```

#Data types
1. logical
2. Character
3. Double: floating point numerical numbers
4. Integer: typeof(7L)

#Special values
1. NA: not available
2. NaN: not a number
3. Inf: positive infinity
4.-Inf: Negative infinity

```{r}
x<-c(1,2,3,4,NA)
mean(x)
mean(x, na.rm = TRUE)
summary(x)

TRUE|NA
FALSE|NA
```
```{r}
# df %>% 
#   pull(y) #pull a vactor from a data frame
```

```{r}
# cat_lovers %>% 
#   mutate(handness=fct_infreq(handness)) %>% #order the bars by frequency count 
#   ggplot(aes(x=handness)) +
#   geom_bar()
```

```{r}
library(tidyverse)
library(skimr)
```
```{r}
hotels <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv", show_col_types = FALSE)
```

```{r}
hotels$arrival_date_month <-factor(hotels$arrival_date_month,
                                   levels = c("January", "February", "March", "April", "May", "June", "July", "August",
                                              "September", "October", "November", "December"))
```


```{r}
hotels %>% 
  filter(is.na(arrival_date_month))
```


```{r fig.height=4.5, fig.width=10}
library(scales)

hotels %>% 
  group_by(hotel, arrival_date_month) %>% 
  summarize(mean_adr=mean(adr)) %>% 
  ggplot(aes(x=arrival_date_month,
             y=mean_adr, 
             group=hotel,
             color=hotel)) +
  geom_line() +
  theme_minimal() +
  labs(x="Arrival Month",
       y="Mean ADR (average daily rate",
       title = "Comparison of resort and city hotel prices across months",
       subtitle = "Resort hotel prices soar in the summer while cith hotel prices remain relatively constant throughout the year",
       color="Hotel type") +
  scale_y_continuous(labels = label_dollar())
```
#Working with dates
```{r}
library(lubridate)
```

```{r}
hotels %>% 
  select(starts_with("arrival_"))
```
```{r}
library(glue)

hotels %>% 
  mutate(
    arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")
  ) %>% 
  select(starts_with("arrival_"))
  
```

```{r}
hotels %>% 
  mutate(
    arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")
  ) %>% 
  count(arrival_date) %>% 
  ggplot(aes(x=arrival_date, y=n, group=1)) +
  geom_line() +
  theme_bw()
```
```{r}
hotels %>% 
  mutate(
    arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))
  ) %>% 
  count(arrival_date) %>% 
  ggplot(aes(x=arrival_date, y=n, group=1)) +
  geom_line() +
  theme_bw()
```

```{r}
hotels %>% 
  mutate(
    arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))
  ) %>% 
  count(arrival_date) %>% 
  ggplot(aes(x=arrival_date, y=n, group=1)) +
 # geom_line() +
  geom_smooth() +
  theme_bw()
```
Importing files

Packages:(readr) (readxl)
Functions
read_csv()-comma delimited
read_csv2()-semicolon separated files
read_tsv()-tab separated files
read_delim()-reads in files with any delimiter
read_fwf()-fixed width files
read_table()-common variation of fixed width files where columns are separated by white space

```{r}
df <- tribble(
  ~x, ~y,
  1, "q",
  2, "b",
  3, "c"
)

write_csv(df,file = "df.csv")
```

```{r}
fs::dir_ls()
```
```{r}
# edi_airbnb <- read_csv("data/edi-airbnb.csv")
# names(edi_airbnb)
```

Remove spaces from column names

Option 1. Define column names
```{r}
# edi_airbnb_col_names <- read_csv("data/edi-airbnb.csv",
#                                  col_names = c("id", "price", "neighbourhood", "accommodates",
#                                                "bathroom", "bedroom", "bed", 
#                                                "review_scores_rating", "n_reviews", "url"))
# names(edi_airbnb)
```

Option 2. Format text to snake_case
```{r}
# edi_airbnb_cleaned_names <- edi_airbnb %>% 
#   janitor::clean_names()
# 
# names(edi_airbnb_cleaned_names) #change to lower case, replace space with underscore
```

NA
```{r}
#read_csv("df-na.csv", na=c(" ", "NA", ".", "9999", "Not applicable"))

#read_csv("df-na.csv",
#col_types = list(col_double, col_character(), col_character()))
```

read rds
```{r}
f<- read_rds("DS4P.rds")
summary(f)
```

read many files
```{r}
#locate the files under a directory
#sales_file<- fs::dir_ls("data/sales")
#sales_file
```

```{r}
library(vroom)
#combine all the files
sales <- vroom::vroom(sales_files, id = "filename")
sales
```

More about writing and reading files
```{r}
library(fs)
gap_tsv<-path_package("gapminder", "extdata", "gapminder.tsv")
gapminder<- read_tsv(gap_tsv)
str(gapminder, give.attr = FALSE)
```
change character to factor
```{r}
gapminder<- gapminder %>% 
  mutate(country = factor(country),
         continent = factor(continent))
str(gapminder)
```
export file
```{r}
gap_life_exp <- gapminder %>% 
  group_by(country, continent) %>% 
  summarize(life_exp = max(lifeExp)) %>% 
  ungroup()

write_csv(gap_life_exp, "gap_life_exp.csv")
```

order levels based on numerical values
```{r}
head(levels(gap_life_exp$country)) #alphabetic order

gap_life_exp <- gap_life_exp %>% 
  mutate(country = fct_reorder(country, life_exp))

head(levels(gap_life_exp$country))
```
Box-Cox transformation
```{r}
library(faraway)
data(butterfat)
```

```{r}
fullmodel = lm(Butterfat ~ Breed * Age, data = butterfat)
plot(fullmodel)
```
```{r}
library(MASS)
bc = boxcox(fullmodel, lambda = seq(-3,3))

#extract best lambda
best.lam = bc$x[which(bc$y == max(bc$y))]

```
```{r}
#boxcox says to transform the data using inverse

fullmodel.inv = lm((Butterfat)^-1 ~ Breed * Age, data = butterfat)
plot(fullmodel.inv)
```

