---
title: "TidyData"
author: "YG"
date: "2023-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Tidy data
1. each variable forms a colum
2. each observation forms a row
3. each type of observational unit forms a table

```{r}
starwars %>%
  select(name, height, mass)
```

```{r}
starwars %>%
  group_by(gender) %>%
  summarize(avg_ht=mean(height, na.rm = T))
```

#Data Structures in R

Matrix: requires all the elements to be of the same type
Data frame: allows for mixed types of variables
Tibble: like a data frame, but works more consistently

A grammar of data wrangling

select: pick columns by name
arrange: reorder rows
slice: pick rows using indexes
filter: pick rows matching criteria
distinct: filter for unique rows
mutate: add new variables
summarize: reduce variables to values
group_by: for grouped operations

#Rules of dplyr
first argument is always a data frame
subsequent argument say what to do with that data frame
always return a data frame
don't modify in place

```{r}
h1<-read_csv("H1.csv")
h1$hotel<"h1"
h2<-read_csv("H2.csv")
h2$hotel<"h2"
hotels<-rbind(h1,h2)
#names(h1)
#names(h2)
```

